### 2015-07-14 V4046Sgr2013/ms_v4046sgr2013_chapter.tex
more changes

These SMA continuum measurements demonstrate that the V4046 Sgr binary joins 
the growing ranks of stars that host massive protoplanetary ``transition" disks 
\citep[e.g.,][]{andrews11}, with a central cavity of radius $\sim$29\,AU (see 
\S \ref{sec:model3}) that is substantially depleted of mm-sized dust particles. 
This classification is commensurate with the SED for V4046 Sgr, shown in Figure 
\ref{fig:summary}(c) and constructed from photometry in the literature 
\citep{hutchinson90,weaver92,jensen96,jensen97,rodriguez10,oberg11}, the 2MASS 
\citep{skrutskie06} and {\it IRAS} \citep{beichmann88} point source catalogs, 
and an archival {\it Spitzer} IRS spectrum.  The SED features the standard 
signature of a transition disk, with a distinctive ``dip" in the continuum near 
10\,$\mu$m suggesting that small, warm dust is preferentially depleted (although
{\it not} absent; see \S\ref{sec:model1}) near the central binary.  An accurate 
determination of the size of this putative cavity from the SED alone is not 
trivial \citep[e.g.,][]{calvet02}, although the very weak excess in the 
near-infrared makes it clear that the dust optical depths must be significantly 
diminished within at least a few AU of the central stars.  Note that the 
observed dip in the continuum covers a much wider wavelength range than was 
inferred by \citet{jensen97}.  Simulations of interactions between a stellar 
binary and its circumbinary disk suggest that the disk material should be 
inwardly truncated at a radius $\sim$2-5$\times$ the binary separation 
\citep[e.g.,][]{artymowicz94}: for V4046 Sgr, the disk truncation should occur 
at $\sim$0.2\,AU, much smaller than the continuum data suggest.  Therefore, the 
observed dust cavity is {\it not} related to interactions with the central 
binary.

@@

While the underlying physical cause of the discrepancy between the size of the 
gas and dust distributions is not clear, there is a 
natural explanation in the growth and inward radial drift of dust particles 
\citep{weidenschilling80,brauer08a,birnstiel10}.  Drift is a fundamental feature
in the migration of disk solids, generated because the radial pressure gradient
in the disk causes the gas to rotate at slightly sub-Keplerian velocities while 
the dust particles are on Keplerian orbits.  The resulting headwind creates a 
drag-force on the dust that relates to the size of the particle, preferentially
moving the large particles towards the central star \citep{weidenschilling77,
takeuchi02}.  As these particles drift inwards, their size distribution is 
continuously evolving as they encounter higher local densities that are 
conducive to growth \citep{birnstiel10}.  The collective effect is to shrink the
radial extent of the dust mass distribution, as the gas disk continues to
viscously spread to larger radii \citep{birnstiel12a}.  Since these processes 
operate on dynamical timescales, this phenomenon should be particularly 
pronounced for the V4046 Sgr disk due to the binarity (increased $M_*$) and 
advanced age of its stellar hosts.  Resolved observations taken across the 
(sub-)millimeter and radio bands may clarify whether the radial grain size 
distribution is consistent with this drift and particle growth scenario 
\citep{guilloteau11,perez12}.

>>>>

These SMA continuum measurements demonstrate that the V4046 Sgr binary joins 
the growing ranks of stars that host massive protoplanetary ``transition" disks 
\citep[e.g.,][]{andrews11}, with a central cavity of radius $\sim$29\,AU (see 
\S \ref{sec:model3}) that is substantially depleted of mm-sized dust particles. 
This classification is commensurate with the SED for V4046 Sgr, shown in Figure 
\ref{fig:summary}(c) and constructed from photometry in the literature 
\citep{hutchinson90,weaver92,jensen96,jensen97,rodriguez10,oberg11}, the 2MASS 
\citep{skrutskie06} and {\it IRAS} \citep{beichman88} point source catalogs, 
and an archival {\it Spitzer} IRS spectrum.  The SED features the standard 
signature of a transition disk, with a distinctive ``dip" in the continuum near 
10\,$\mu$m suggesting that small, warm dust is preferentially depleted (although
{\it not} absent; see \S\ref{sec:model1}) near the central binary.  An accurate 
determination of the size of this putative cavity from the SED alone is not 
trivial \citep[e.g.,][]{calvet02}, although the very weak excess in the 
near-infrared makes it clear that the dust optical depths must be significantly 
diminished within at least a few AU of the central stars.  Note that the 
observed dip in the continuum covers a much wider wavelength range than was 
inferred by \citet{jensen97}.  Simulations of interactions between a stellar 
binary and its circumbinary disk suggest that the disk material should be 
inwardly truncated at a radius $\sim$2-5$\times$ the binary separation 
\citep[e.g.,][]{artymowicz94}: for V4046 Sgr, the disk truncation should occur 
at $\sim$0.2\,AU, much smaller than the continuum data suggest.  Therefore, the 
observed dust cavity is {\it not} related to interactions with the central 
binary.

@@

While the underlying physical cause of the discrepancy between the size of the 
gas and dust distributions is not clear, there is a 
natural explanation in the growth and inward radial drift of dust particles 
\citep{weidenschilling80,brauer08,birnstiel10}.  Drift is a fundamental feature
in the migration of disk solids, generated because the radial pressure gradient
in the disk causes the gas to rotate at slightly sub-Keplerian velocities while 
the dust particles are on Keplerian orbits.  The resulting headwind creates a 
drag-force on the dust that relates to the size of the particle, preferentially
moving the large particles towards the central star \citep{weidenschilling77b,
takeuchi02}.  As these particles drift inwards, their size distribution is 
continuously evolving as they encounter higher local densities that are 
conducive to growth \citep{birnstiel10}.  The collective effect is to shrink the
radial extent of the dust mass distribution, as the gas disk continues to
viscously spread to larger radii \citep{birnstiel12a}.  Since these processes 
operate on dynamical timescales, this phenomenon should be particularly 
pronounced for the V4046 Sgr disk due to the binarity (increased $M_*$) and 
advanced age of its stellar hosts.  Resolved observations taken across the 
(sub-)millimeter and radio bands may clarify whether the radial grain size 
distribution is consistent with this drift and particle growth scenario 
\citep{guilloteau11,perez12}.

### 2015-07-14 ms_radial_092713/ms_radial_chapter.tex
more changes

A subset of T Tauri and Herbig Ae disks have large, mass-depleted cavities in 
their inner regions, inferred from their infrared spectra 
\citep[e.g.,][]{Kim13} or spatially resolved imaging at infrared 
\citep[e.g.,][]{Geers2007} and mm wavelengths \citep[e.g.,][]{Brown09}.  The
cavities in these ``transition'' disks have radii ranging from $\sim$4\,AU 
\citep[TW Hya;][]{Calvet02,Hughes07} to 20\,AU \citep[GM Aur;][]{hughes09} to 
40--50\,AU \citep[LkCa 15;][]{Andrews11,Isella12} to 140\,AU \citep[HD 
142527;][]{casassus13}.  The suspicion is that dynamical interactions with 
nascent planetary systems evacuate the inner disk regions 
\citep[e.g.,][]{Zhu11}.

@@

The role of stars in galaxies might be played by planets in transition disks.  
We like this interpretation, but it is not without difficulty.  The cavities 
are so large (they are not annular gaps) and so clean that appeals are made to 
rather extreme parameters.  As many as four planets, each weighing up to 
$\sim$10 $M_{\rm Jup}$, have been posited to carve out a given cavity 
\citep{Zhu11,Dodson11}.\footnote{Even multiple super-Jovian planets are 
reportedly not enough: grain growth and dust filtration at the cavity rim are 
invoked in addition to planets to make the cavity sufficiently transparent 
\citep{Zhu11,Dong12}.}  At least in the case of the transition disk around LkCa 
15, there might be one (but so far only one) actively accreting super-Jovian 
planet located well inside the disk cavity \citep{Kraus12}.  Some assistance in 
driving the accretion flow inside the cavity may be obtained from the 
gravitational potential of the outer disk itself, if that disk is sufficiently 
clumpy (for examples of clumpy disks, see IRS 48 or HD 142527; 
\citealt{vandermarel13} and \citealt{casassus13}, respectively).  Just inside a 
non-axisymmetric outer disk, gas streamlines can cross, shock, and plunge 
inward \citep[e.g.,][]{pac77}.

Disk radial velocities driven by embedded massive planets or brown dwarfs can 
be healthy fractions of orbital (azimuthal) velocities.  We derive a crude 
estimate\footnote{Our ``zero-dimensional'' analysis ignores the fact that 
actual disk flows around planets are two-dimensional and dominated by
non-axisymmetric streams flowing in and around planetary Hill spheres.
Still, some indirect empirical support for the scalings in equation 
(\ref{eqn:simple}) has been found by \cite{Fung13}.
}
 for $v_R$ by equating the one-sided Lindblad torque --- the amount of 
angular momentum transmitted outward across a planet's orbit 
\citep{goldreich80} --- with the angular momentum advected inward by accretion, 
$\dot{M} v_{\rm K} R$.  This balance yields
\begin{equation}
v_R \sim \left( \frac{M_p}{M_\ast} \right)^2 \left( \frac{H}{R} \right)^3 v_{\rm K} \label{eqn:simple}
\end{equation}
where we have used the total linear Lindblad torque, integrated over all 
resonances outside the torque cut-off zone at distances $\gtrsim H$ away from 
the planet (see, e.g., \citealt{goodman01} or \citealt{crida06}, and references 
therein).  For a planet mass $M_p \sim 10\,M_{\rm Jup}$, a host stellar mass 
$M_\ast \sim 1\,M_\odot$, and a typical disk aspect ratio $H/R \sim 0.1$, 
equation (\ref{eqn:simple}) evaluates to
\begin{equation}
v_R \sim 0.1 \,v_{\rm K} \sim 0.5 \left( \frac{30\,{\rm AU}}{R} \right)^{1/2}\,{\rm km} \,\, {\rm s}^{-1}. 
\label{eqn:simple1}
\end{equation}
Modern heterodyne receivers on (sub)millimeter wavelength interferometers 
routinely observe disks at $\sim$0.1 km s$^{-1}$ resolution, and the Atacama 
Large Millimeter Array (ALMA) is expected to provide high sensitivity data at 
finer velocity scales yet.  Evidence for free-fall radial velocities can be 
found in the HCO$^+$ filament within the disk cavity around HD 142527.  
\citet{casassus13} discovered this filament to be oriented nearly radially; in 
estimating $\dot{M}$ for this system, they assumed $v_R \sim v_{\rm K}$ 
(although the velocities of the HCO$^+$ filament were not actually quoted in
their study).

@@

We compute an analogous 3D disk model that simulates the $^{12}$CO $J$=3$-$2 
line emission from a disk with a total gas mass of 0.01\,$M_\odot$, a spatially 
constant CO:H$_2$ number-abundance fraction of $10^{-4}$, and a Gaussian 
vertical density distribution ($\exp(-z^2/2H^2)$ with a scale height of 
$H=0.1R$).  The surface density profile is that of a thin, viscously accreting 
disk \citep{bellpringle74}: $\Sigma\propto R^{-1}\exp(-R/R_s)$ with scale 
radius $R_s=50$ AU and an outer truncation radius of 300 AU.  Since our focus 
in this paper is on the velocity field and not on the density structure, we 
have not included any reduction of $\Sigma$ within the cavity.  This 
simplification should not introduce serious error as long as the cavity region 
remains optically thick in $^{12}$CO emission, as is apparently the case in 
reality \citep{bruderer13}.  The model is vertically isothermal with a 
power-law radial temperature profile: $T(R) = 25\,(R/100$\,AU$)^{-0.5}$\,K.  
The mass of the central star is 1\,$M_\odot$, the distance to the observer is 
145\,pc, and the cavity radius is $R_{\rm cav} = 50$\,AU.  The disk inclination
$i$ and radial inflow parameter $\chi$ are variable.  We use the modeling 
capabilities of {\tt LIME} \citep{brinch10} to simulate high-resolution channel 
maps assuming LTE conditions and a turbulent linewidth of $\xi = 
0.01$\,km\,s$^{-1}$.  We then use {\tt CASA} to sample and deconvolve the 
Hanning-smoothed visibilities with a spectral resolution of 0.212\,km\,s$^{-1}$ 
and a spatial resolution of 0\farcs1.  The right panel of Figure 
\ref{fig:models} shows the channel maps from a 3D disk model ($\chi=0$), zoomed 
into the inner $\pm$0\farcs5.  Comparing the left and right panels of Figure 
\ref{fig:models} reveals how the velocity field of the razor-thin model offers 
a guide to the morphology of line emission in the channel maps of the 3D model 
\citep[see also][]{beckwith93}.

>>>>

A subset of T Tauri and Herbig Ae disks have large, mass-depleted cavities in 
their inner regions, inferred from their infrared spectra 
\citep[e.g.,][]{Kim13} or spatially resolved imaging at infrared 
\citep[e.g.,][]{geers2007} and mm wavelengths \citep[e.g.,][]{brown09}.  The
cavities in these ``transition'' disks have radii ranging from $\sim$4\,AU 
\citep[TW Hya;][]{calvet02,hughes07} to 20\,AU \citep[GM Aur;][]{hughes09} to 
40--50\,AU \citep[LkCa 15;][]{andrews11b,isella12} to 140\,AU \citep[HD 
142527;][]{casassus13}.  The suspicion is that dynamical interactions with 
nascent planetary systems evacuate the inner disk regions 
\citep[e.g.,][]{Zhu11}.

@@

The role of stars in galaxies might be played by planets in transition disks.  
We like this interpretation, but it is not without difficulty.  The cavities 
are so large (they are not annular gaps) and so clean that appeals are made to 
rather extreme parameters.  As many as four planets, each weighing up to 
$\sim$10 $M_{\rm Jup}$, have been posited to carve out a given cavity 
\citep{Zhu11,dodson11}.\footnote{Even multiple super-Jovian planets are 
reportedly not enough: grain growth and dust filtration at the cavity rim are 
invoked in addition to planets to make the cavity sufficiently transparent 
\citep{Zhu11,dong12}.}  At least in the case of the transition disk around LkCa 
15, there might be one (but so far only one) actively accreting super-Jovian 
planet located well inside the disk cavity \citep{Kraus12}.  Some assistance in 
driving the accretion flow inside the cavity may be obtained from the 
gravitational potential of the outer disk itself, if that disk is sufficiently 
clumpy (for examples of clumpy disks, see IRS 48 or HD 142527; 
\citealt{vandermarel13} and \citealt{casassus13}, respectively).  Just inside a 
non-axisymmetric outer disk, gas streamlines can cross, shock, and plunge 
inward \citep[e.g.,][]{paczynski77}.

Disk radial velocities driven by embedded massive planets or brown dwarfs can 
be healthy fractions of orbital (azimuthal) velocities.  We derive a crude 
estimate\footnote{Our ``zero-dimensional'' analysis ignores the fact that 
actual disk flows around planets are two-dimensional and dominated by
non-axisymmetric streams flowing in and around planetary Hill spheres.
Still, some indirect empirical support for the scalings in equation 
(\ref{eqn:simple}) has been found by \cite{fung13}.
}
 for $v_R$ by equating the one-sided Lindblad torque --- the amount of 
angular momentum transmitted outward across a planet's orbit 
\citep{goldreich80} --- with the angular momentum advected inward by accretion, 
$\dot{M} v_{\rm K} R$.  This balance yields
\begin{equation}
v_R \sim \left( \frac{M_p}{M_\ast} \right)^2 \left( \frac{H}{R} \right)^3 v_{\rm K} \label{eqn:simple}
\end{equation}
where we have used the total linear Lindblad torque, integrated over all 
resonances outside the torque cut-off zone at distances $\gtrsim H$ away from 
the planet (see, e.g., \citealt{goodman01} or \citealt{crida06}, and references 
therein).  For a planet mass $M_p \sim 10\,M_{\rm Jup}$, a host stellar mass 
$M_\ast \sim 1\,M_\odot$, and a typical disk aspect ratio $H/R \sim 0.1$, 
equation (\ref{eqn:simple}) evaluates to
\begin{equation}
v_R \sim 0.1 \,v_{\rm K} \sim 0.5 \left( \frac{30\,{\rm AU}}{R} \right)^{1/2}\,{\rm km} \,\, {\rm s}^{-1}. 
\label{eqn:simple1}
\end{equation}
Modern heterodyne receivers on (sub)millimeter wavelength interferometers 
routinely observe disks at $\sim$0.1 km s$^{-1}$ resolution, and the Atacama 
Large Millimeter Array (ALMA) is expected to provide high sensitivity data at 
finer velocity scales yet.  Evidence for free-fall radial velocities can be 
found in the HCO$^+$ filament within the disk cavity around HD 142527.  
\citet{casassus13} discovered this filament to be oriented nearly radially; in 
estimating $\dot{M}$ for this system, they assumed $v_R \sim v_{\rm K}$ 
(although the velocities of the HCO$^+$ filament were not actually quoted in
their study).

@@

We compute an analogous 3D disk model that simulates the $^{12}$CO $J$=3$-$2 
line emission from a disk with a total gas mass of 0.01\,$M_\odot$, a spatially 
constant CO:H$_2$ number-abundance fraction of $10^{-4}$, and a Gaussian 
vertical density distribution ($\exp(-z^2/2H^2)$ with a scale height of 
$H=0.1R$).  The surface density profile is that of a thin, viscously accreting 
disk \citep{lynden-bell74}: $\Sigma\propto R^{-1}\exp(-R/R_s)$ with scale 
radius $R_s=50$ AU and an outer truncation radius of 300 AU.  Since our focus 
in this paper is on the velocity field and not on the density structure, we 
have not included any reduction of $\Sigma$ within the cavity.  This 
simplification should not introduce serious error as long as the cavity region 
remains optically thick in $^{12}$CO emission, as is apparently the case in 
reality \citep{bruderer13}.  The model is vertically isothermal with a 
power-law radial temperature profile: $T(R) = 25\,(R/100$\,AU$)^{-0.5}$\,K.  
The mass of the central star is 1\,$M_\odot$, the distance to the observer is 
145\,pc, and the cavity radius is $R_{\rm cav} = 50$\,AU.  The disk inclination
$i$ and radial inflow parameter $\chi$ are variable.  We use the modeling 
capabilities of {\tt LIME} \citep{brinch10} to simulate high-resolution channel 
maps assuming LTE conditions and a turbulent linewidth of $\xi = 
0.01$\,km\,s$^{-1}$.  We then use {\tt CASA} to sample and deconvolve the 
Hanning-smoothed visibilities with a spectral resolution of 0.212\,km\,s$^{-1}$ 
and a spatial resolution of 0\farcs1.  The right panel of Figure 
\ref{fig:models} shows the channel maps from a 3D disk model ($\chi=0$), zoomed 
into the inner $\pm$0\farcs5.  Comparing the left and right panels of Figure 
\ref{fig:models} reveals how the velocity field of the razor-thin model offers 
a guide to the morphology of line emission in the channel maps of the 3D model 
\citep[see also][]{beckwith93}.
